---
title: "CDI app files preparation"
author: "Magdalena Krysztofiak"
date: "2026-01-23"
output: html_document
---

```{r setup}
library(tidyverse)
```

## Norwegian WG

### Items

```{r import-wg}
read_csv("Data/norwegian_wg_gest_items.csv") -> no_wg_gest
read_csv("Data/norwegian_wg_comp_items.csv") -> no_wg_comp
read_csv("Data/norwegian_wg_prod_items.csv") -> no_wg_prod
#bind
bind_rows(no_wg_gest, no_wg_comp, no_wg_prod) -> no_wg
```

First, we need to get semantic category for each item in Norwegian
This can be extracted from the translation file...
```{r descriptive-categories-wg}
read_csv2("Items/no/wg/translations.csv") %>% 
  filter(text_type == "catHeader") %>%
  select(-text_type) %>%
  rename(category_descriptive = text) -> no_wg_categories

read_csv2("Items/no/wg/items.csv") %>% 
  select(item_id, definition, category, type) %>%
  mutate(category = if_else(is.na(category),
                            type,category)) %>%
  mutate(type = if_else(grepl("gestures_*",type),
                        "gestures",type)) %>%
  left_join(no_wg_categories, 
            by = c("category", "type" = "item_type")) -> no_wg_items_categories
```

Secondly, we merge the categories with the CAT items
Some items have missing category bc. duplicates (vann) are stored differently
in each file.
```{r join-cat-items-with-category}
no_wg %>% left_join(no_wg_items_categories, 
                    by = c("id" = "item_id", 
                           "item" = "definition")) -> no_wg

#check and save items with missing information on category for manual coding
#no_wg %>% filter(is.na(category)) %>% 
#  select(id, item, category, type, category_descriptive) %>%
#  distinct() %>%
#  mutate(updated_item = "") %>%
#  write_csv("no_missing_category.csv")

#import missing categories
read_csv2("no_wg_filled_missing_category.csv") %>% 
  right_join(no_wg, by = c("id", "item"), suffix = c("_x", "_y")) %>% 
  mutate(
  category = coalesce(category_y, category_x),
  type = coalesce(type_y, type_x),
  category_descriptive = coalesce(category_descriptive_y, category_descriptive_x),
  item = coalesce(updated_item, item)
) %>% select(-ends_with("_x"), -ends_with("_y"), -updated_item) -> no_wg

#check if there are any NA values
#sum(is.na(no_wg))
```

Third, we merge categories with question.
```{r no-wg-question}
no_wg %>% 
  mutate(question_ = case_when(group == "Prod" ~ "Sier barnet…",
                               group == "Comp" ~ "Forstår barnet…",
                               group == "Gest" ~ "")) %>%
  unite("question", category_descriptive, question_, sep = ": ") %>%
  select(a1,d,item,question,group,id) %>% write_csv("CAT batch files/no/wg/items.csv")
```

### StartThetas

We need to bind files with StartThetas for gestures, comprehension, production

```{r no-wg-startthetas}
rbind(read_csv("Data/norwegian_wg_gest_startThetas.csv"),
      read_csv("Data/norwegian_wg_comp_startThetas.csv"),
      read_csv("Data/norwegian_wg_prod_startThetas.csv")) -> no_wg_thetas
write_csv(no_wg_thetas,"CAT batch files/no/wg/startThetas.csv")
```

## Norwegian WS

```{r import-ws}
read_csv("Data/norwegian_ws_items.csv") -> no_ws
```

First, we need to get semantic category for each item in Norwegian
This can be extracted from the translation file...
```{r descriptive-categories-ws}
read_csv2("Items/no/ws/translations.csv") %>% 
  filter(text_type == "catHeader") %>%
  select(-text_type) %>%
  rename(category_descriptive = text) -> no_categories

read_csv2("Items/no/ws/items.csv") %>% 
  select(item_id, definition, category, type) %>%
  mutate(category = if_else(is.na(category),
                            type,category)) %>%
  mutate(type = if_else(grepl("gestures_*",type),
                        "gestures",type)) %>%
  left_join(no_categories, 
            by = c("category", "type" = "item_type")) -> no_items_categories
```

Secondly, we merge the categories with the CAT items
Some items have missing category bc. duplicates (vann) are stored differently
in each file.
```{r join-cat-items-with-category-ws}
no_ws %>% left_join(no_items_categories, 
                    by = c("id" = "item_id", 
                           "item" = "definition")) -> no_ws

#check and save items with missing information on category for manual coding
#no_ws %>% filter(is.na(category)) %>% 
#  select(id, item, category, type, category_descriptive) %>%
#  distinct() %>%
#  mutate(updated_item = "") %>%
#  write_csv("no_ws_missing_category.csv")

#import missing categories
read_csv2("no_ws_filled_missing_category.csv") %>% 
  right_join(no_ws, by = c("id", "item"), suffix = c("_x", "_y")) %>% 
  mutate(
  category = coalesce(category_y, category_x),
  type = coalesce(type_y, type_x),
  category_descriptive = coalesce(category_descriptive_y, category_descriptive_x),
  item = coalesce(updated_item, item)
) %>% select(-ends_with("_x"), -ends_with("_y"), -updated_item) -> no_ws
```

Third, we merge categories with question and write.
```{r no-ws-question}
no_ws %>% 
  mutate(question_ =  "Sier barnet…") %>%
  unite("question", category_descriptive, question_, sep = ": ") %>%
  select(a1,d,item,question,group,id) %>% write_csv("CAT batch files/no/ws/items.csv")
```

