---
title: "IRT model fitting"
subtitle: "Norwegian CDI:WG Comprehension"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    code_folding: show
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE,
                      echo = FALSE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = "center",
                      fig.width = 6,
                      fig.height = 4)
```

# Setup

Below are the settings for the analyses reported in this document.

```{r settings, echo = TRUE}
language <- "Norwegian"
scale <- "WG Comprehension" # one of "WG Comprehension", "WG Production", "WG Gestures", "WS"
```

Plot titles, file names etc. are based on them.

```{r analyses_setup}
# Additional setup:
functions_dir <- "Functions_new" # "Functions_new" keeps updated function definition
data_dir <- "Data"

version <- strsplit(scale, " ")[[1]]
subscale <- "Prod"
if(length(version) > 1) substr(version[2], 1, 4) -> subscale
version <- version[1]
if(version == "WS") {
  filename_prefix <- tolower(paste(language, version, sep="_"))
} else {
  filename_prefix <- tolower(paste(language, version, subscale, sep="_"))
}
filename_prefix <- paste(data_dir, filename_prefix, sep="/")
filename_data <- paste(data_dir, tolower(paste(language, version, sep="_")), sep="/")
```

The following  libraries are loaded.

```{r libraries, echo=TRUE, results='hide'}
library(beepr)
library(tidyverse)
library(mirt)
```

Functions from "`r functions_dir`" directory are sourced.
Data are expected and saved in "`r data_dir`" directory.
 
```{r functions}
files <- list.files(path = functions_dir, pattern = "\\.R$", full.names = TRUE)
for (f in files) source(f)
```


## Data preparation

Data may be downloaded from the Wordbank database and saved locally
for quicker access in the future. See and run the chunk below for that
(modifying calls to `select()` if more data, demographics in particular, is needed).

```{r wordbank_download, eval=FALSE, echo=TRUE}
# Get responses and item data:
wordbankr::get_instrument_data(language, version,
                               administration_info=TRUE,
                               item_info=TRUE) %>%
     filter(is_norming) -> responses
# More item data available, if needed -- modify select():
responses %>%
     select(child_id, item_id, item_definition, english_gloss, item_kind, value) %>%
     write_csv(paste(filename_data, "responses.csv", sep="_"))

# Get demographics data:
wordbankr::get_administration_data(language, version,
                                   include_demographic_info=TRUE,
                                   include_birth_info=TRUE,
                                   include_health_conditions=TRUE,
                                   include_language_exposure=TRUE,
                                   include_study_internal_id=TRUE) %>%
     filter(is_norming) -> participants
# More demographics data available, if needed -- modify select():
participants %>%
     select(child_id, age, sex) %>%
     write_csv(paste(filename_data, "participants.csv", sep="_"))
```

There should be two *csv* files:

1. "`r paste(filename_data, "responses.csv", sep="_")`" and
2. "`r paste(filename_data, "participants.csv", sep="_")`".

First one is used to extract the list of `r paste(language, scale)` items
(with their definitions and English glosses, if available).
Relevant responses are recoded as 0's and 1's appropriately.

Second one contains demographic data (*age* and *sex* in particular).

```{r data_import}
# Read csv files:
read_csv(paste(filename_data, "participants.csv", sep="_")) -> participants
read_csv(paste(filename_data, "responses.csv", sep="_")) -> responses

# Filter & recode responses based on scale:
if(subscale == "Gest") {
     responses %>% filter(str_starts(item_kind, "gestures")) %>%
          mutate(value = case_match(value,
                                   c("not yet", "no") ~ 0,
                                   c("often", "sometimes", "yes") ~ 1,
                                   .default = 0)) -> responses
} else {
     if(subscale == "Prod") {
          positive <- c("produces", "understands")
     } else {
          positive <- "understands"
     }
     responses %>% filter(item_kind == "word") %>%
          mutate(value = case_match(value,
                                    positive ~ 1,
                                    NA ~ 0,
                                    .default = 0)) -> responses
}

# Extract instrument contents (item definitions, English glosses, and IDs):
responses %>% distinct(item_definition, english_gloss, item_id) -> cdi
```

There are three key objects needed for all the analyses done and presented here:

1. `responses` matrix (administrations x items; 0/1)
2. `cdi` data frame (items with their definitions & glosses)
3. `scores` data frame (participant information: raw scores & demographics)

To prepare the `responses` matrix we eliminate *floor* and *ceiling* (zero variance) items.

**NB:** `Mirt` documentation suggests removing, as *floor* and *ceiling*,
items with difficulties > *.95* and < *.05*. **Check and explore in the future?**

```{r data_processing}
# Filter out floor & ceiling items:
original_n <- length(unique(responses$child_id))
responses %>% group_by(item_id) %>% filter(var(value) > 0) %>%
     pivot_wider(names_from = "item_id", values_from = "value", id_cols = "child_id") -> responses
# Remove them from the instrument data:
original_cdi_n <- nrow(cdi)
cdi %>% filter(item_id %in% colnames(responses)) -> cdi

# Calculate scores and merge with demographics:
responses %>% rowwise() %>% mutate(score = sum(across(starts_with("item_")))) %>%
     select(1, last_col()) %>% left_join(participants) -> scores

# Create responses matrix:
responses <- as.matrix(select(responses, starts_with("item_")))
```

There are `r original_n` `r paste(language, scale)` administrations
and `r original_cdi_n` `r paste(language, scale)` items.
After removing *floor* and *ceiling* items there are
`r nrow(responses)` administrations and `r ncol(responses)` items
for IRT model fitting.

```{r scores_exporting}
scores %>% write_csv(paste(filename_prefix, "scores.csv", sep="_"))
```

We export `scores` data frame to "`r paste(filename_prefix, "scores.csv", sep="_")`"
to use it once we have a model.


# Number of quadrature points

## Introduction

The number of quadrature points (`quadpts` parameter in `mirt` package)
is a parameter important for numerical integration used
in the IRT model fitting. We know that for some of our datasets the default number
results in the model not converging. Therefore with new data we used to
routinely first try model fitting with an increasing number of `quadpts`,
starting with the default 61, and going up by an arbitrary number of arbitrary
steps. We would then compare the convergence of the model, its fit, and
the parameter estimates for the first item, and choose the `quadpts`
for which these statistics stabilised (roughly).

Given that model fitting is a one-off thing (we save the model for later reuse),
it seems more convenient -- and also beneficial in the context of cross data comparisons --
to set `quadpts` reasonably high and keep it constant. The following section
is totally optional.

## Model comparisons (*only if needed*)

From time to time we might want to run these comparisons. E.g., if new analyses
yield results different from the old ones, or for new data, especially if
the model doesn't converge on the first try.

The `quadpts` values tested and the choice criteria are not precise but
it shouldn't matter. The greater the value the longer fitting time but greater precision.
The default, historical, vector `quadpts = c(61, 151, 301, 451, 601, 751)` can be
changed as required.

```{r quadpts_comp, eval=FALSE, echo=TRUE}
quadpts <- quadpts_comp(responses, quadpts = c(61, 151, 301, 451, 601, 751), NCYCLES = 100000,
             output_file = paste(filename_prefix, "quadpts.RData", sep="_"))
```

Results of the above code are saved to a file
("`r paste(filename_prefix, "quadpts.RData", sep="_")`"), so it needs to be run
only once.

The convergence status and model fits (logLik) for different values of `quadpts`
(which can be printed below) are read from that file.

```{r quadpts_comp_fit, eval=FALSE}
load(paste(filename_prefix, "quadpts.RData", sep="_"))
quadpts[1:3,]
```

The same applies to parameter estimates for the first item.

```{r quadpts_comp_estimates, eval=FALSE}
sapply(1:ncol(quadpts), function(x) quadpts[,x]$item1pars)
```


## Setting

Based on the above (**or not**) we set the number of quadrature points.

```{r quadpts_set, echo=TRUE}
quadtp <- 600
```


```{r fitting_iteration, echo=TRUE}
fitting_iteration <- 1
# Model fitting

Our fitting the IRT model is an iterative process that involves eliminating items and
refitting the model in two loops, one within the other.

1. In the "inner" loop we iteratively fit the model with `r quadtp` quadrature points,
remove potential misfits (significant S-X²) and refit until we have fits for all items.

2. Then we examine *local dependency* (Cramer's $V$) and, if there are items to remove,
we go back to point 1 (this is the "outer" loop).

The "inner" loop, given its simple and easy to set criterion (significance), is fully automatic
and built into a custom function from "`r functions_dir`" directory. The "outer" loop is
run manually. With the current criterion (Cramer's $V$ threshold) it could be made automatic
as well but, given that this approach is likely suboptimal ([see below](more-on-the-issue)),
we keep it as it is: if there are items to be removed in step 2, the code execution halts and
rerunning previous chunks is suggested.

Key intermediate results are saved to files in "`r data_dir`" directory with
`r paste(language, scale)` and "outer" loop iteration number kept in filenames.
If the document knits, the whole model fitting procedure has already succeeded and
the summary is simply printed below.

```


## Step 1. Misfits removal

First we repeatedly fit the model with `r quadtp` quadrature points,
each time remove any misfits until we have fits for all items.
The final model and the list of removed items are saved to
"`r paste(filename_prefix, fitting_iteration, "model.RData", sep="_")`" file.

```{r misfits_removal, eval=FALSE}
model <- misfits_removal(responses=responses, quadtps=quadtp, NCYCLES=100000, p=0.001,
                          output_file = paste(filename_prefix, fitting_iteration, "model.RData", sep="_"))
```

We print the model's basic summary.

```{r model_summary}
load(paste(filename_prefix, fitting_iteration, "model.RData", sep="_"))
mod <- model[[1]]
items_removed <- model[[2]]
item_fit <- model[[3]]
rm(model) # to save space

mod
```

`r length(unlist(items_removed))` items were misfits in the process of model fitting.

```{r items_removed_list}
cdi %>% filter(item_id %in% unlist(items_removed))
```

They were removed in `r length(items_removed)` iterations before a model with no misfits was obtained
(misfits removing function iterations, not to be confused with `fitting_iteration`
equaled `r fitting_iteration`).

```{r items_removed}
if(length(items_removed)) {
  responses <- responses[, -which((colnames(responses) %in% unlist(items_removed)))]
  cdi %>% filter(! item_id %in% unlist(items_removed)) -> cdi
  if (nrow(cdi) == ncol(responses)) print(paste0(ncol(responses), " items left after removal."))
}
```


## Step 2. Local dependence

As the next step we check pairwise correlations between
items after taking the model (i.e., the underlying trait) into account.
For each pair of residuals we calculate Cramer's $V$, a measure of
association between two dichotomous variables.

### More on the issue

What `mirt` package returns are **signed** coefficients and the sign indicates the direction of
the relation. Positive relation (*local dependence*) means that two items
are more related than the model predicts, i.e., are too similar or redundant.
IRT assumes *local independence*: after conditioning on the latent trait $\theta$,
responses are assumed to be statistically independent; or in plain English,
if we know a person’s ability level, knowing their response to one item
should not help us predict their response to another; *Chat GPT* assistance).

Keeping locally dependent items in the item pool may lead to a number of
distortions in CAT measurement (crucially, among other, inflated $\theta$ estimates;
*Chat GPT* assistance). This is why it is important to monitor this problem.
As with any correlation metric however, its interpretation is always arbitrary.

So far we have simply removed all item pairs with $V > .5$.
However, $.5$ might be too *high* as a threshold, potentially leaving in too many
locally dependent items. At the same time, removing *both items* is obviously
too harsh, reducing the item pool excessively,
since the problem is the relation between them, not the items themselves.

Deciding on which item to keep is not straightforward though.
The simplest option is to rely on a single statistic, e.g., *item's fit*,
but high $V$ pairs are often *gender pairs* like "mommy--daddy", "boy--girl",
"grandpa--granny", "his--her": keeping only one seems somewhat questionable
(this is in fact a potential guideline for future CDI development/revisions:
**make such pairs single items**).
Moreover, the same item may appear in a number of high $V$ pairs:
it's reasonable to remove that item, even if its fit is better than the other items'.

Finally, so far we have overlooked negative $V$'s:
they suggest the violation of the local independence assumption as well
(hidden multidimensionality might be at play).
However, they seem to be less common and might be less distortive for CAT measurement
(*Chat GPT* assistance; further checks needed).

All the above suggests that diagnosing local dependence is a complex issue that
might require careful examination and deliberate decisions, rether than
a purely algorithmic procedure.


We calculate Cramer's $V$ for each pair of residuals and save the results to
"`r paste(filename_prefix, fitting_iteration, "residuals.RData", sep="_")`" file.
### Summary & selection

```{r local_dependence, eval=FALSE, results='hide'}
# I can't make the following code hide the results...
residuals <- residuals(mod, df.p=FALSE) # TODO: explore other parameters?
save(residuals, file = paste(filename_prefix, fitting_iteration, "residuals.RData", sep="_"))
```

The histogram below shows the (**signed**) Cramer's $V$ distribution.

```{r local_dependence_histogram}
load(paste(filename_prefix, fitting_iteration, "residuals.RData", sep="_"))
residuals_up <- residuals
residuals_up[lower.tri(residuals_up)] <- NA # upper diagonal contains signed Cramers V coefficients
hist(residuals_up, main="Histogram of signed Cramer's V coefficients", xlab= "Cramer's V")
```

These are the descriptive statistics of the distribution.

```{r local_dependence_summary}
summary(as.vector(residuals_up))[-7]
# -7 to avoid reporting NAs, which would be confusing as lower triangle is filled with them
```

To display *questionably* mutually dependent items
we use $V > .3$ as the threshold. We print their definitions, English glosses,
and fit statistics (\mathrm{S\!-\!X}^2).

```{r local_dependence_positive}
cramers_V_df(arrayInd(which(residuals_up > 0.3), dim(residuals)), cdi, item_fit) %>%
  select(- starts_with("item_id"))
```

In the same way we also print all pairs with *questionably* high negative
values ($V < -.3$).

```{r local_dependence_negative}
cramers_V_df(arrayInd(which(residuals_up < -0.3), dim(residuals)), cdi, item_fit) %>%
  select(- starts_with("item_id"))
```


To **remove** the most highly mutually dependent items
we use $V > .5$ as the threshold.

```{r local_dependence_remove}
indexes <- arrayInd(which(residuals_up > 0.5), dim(residuals))
cramers_V_df(indexes, cdi, item_fit) %>%
  select(- starts_with("item_id"))
unique(as.vector(indexes)) -> items_to_remove_ids
if(length(items_to_remove_ids)) {
     responses[, -items_to_remove_ids] -> responses
     cdi[-items_to_remove_ids,] -> cdi
}
```

Number of removed items is `r length(items_to_remove_ids)`.

If it's **greater than zero** we need to increase the `fitting_iteration` number
by one (see chunk below) and repeat everything from
the [Misfits removal](#misfits-removal) section onwards
(also don't forget to update the `fitting_iteration` chunk right before
[Misfits removal](#misfits-removal)).

```{r fitting_iteration_increment, eval=FALSE, echo=TRUE}
fitting_iteration <- fitting_iteration + 1
```

If number of removed items is zero we're done.
The final data frame of `r paste(language, scale)` items retained
after model fitting is exported for future use to
`r paste(filename_prefix, fitting_iteration, "items.csv", sep="_")`

```{r items_exporting}
cdi %>% write_csv(paste(filename_prefix, fitting_iteration, "items.csv", sep="_"))
```

