---
title: "Norwegian CDI:WG Gestures analysis"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE,
                      echo = FALSE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = "center",
                      fig.width = 6,
                      fig.height = 4)
```

# Setup

Below are the settings for the analyses reported in this document.

```{r settings, echo = TRUE}
functions_dir <- "Functions_new" # "Functions_new" keeps updated function definition
data_dir <- "Data"
language <- "Norwegian"
scale <- "WG Gestures" # one of "WG Comprehension", "WG Production", "WG Gestures", "WS"
```

Plot titles, file names etc. are based on them.

```{r analyses_setup}
version <- strsplit(scale, " ")[[1]]
subscale <- "Prod"
if(length(version) > 1) substr(version[2], 1, 4) -> subscale
version <- version[1]
if(version == "WS") {
  filename_prefix <- tolower(paste(language, version, sep="_"))
} else {
  filename_prefix <- tolower(paste(language, version, subscale, sep="_"))
}
filename_prefix <- paste(data_dir, filename_prefix, sep="/")
filename_data <- paste(data_dir, tolower(paste(language, version, sep="_")), sep="/")
```

The following  libraries are loaded.

```{r libraries, echo=TRUE, results='hide'}
library(beepr)
library(wordbankr)
library(tidyverse)
library(mirt)
library(mirtCAT)
library(parallel)
library(ggpubr)
```

Functions from `r paste0("`", functions_dir, "`")` directory are sourced.
 
```{r functions}
files <- list.files(path = functions_dir, pattern = "\\.R$", full.names = TRUE)
for (f in files) source(f)
```


# Data preparation

Data may be downloaded from the Wordbank database and saved locally
for quicker access in the future. See and run the chunk below for that
(modifying calls to `select()` if more data, demographics in particular, is needed)

```{r wordbank_download, eval=FALSE, echo=TRUE}
# Get responses and item data:
get_instrument_data(language, version,
                    administration_info=TRUE,
                    item_info=TRUE) %>%
                    filter(is_norming) -> responses
# More item data available, if needed -- modify select():
responses %>%
  select(child_id, item_id, item_definition, english_gloss, item_kind, value) %>%
  write_csv(paste(filename_data, "responses.csv", sep="_"))

# Get demographics data:
get_administration_data(language, version,
                        include_demographic_info=TRUE,
                        include_birth_info=TRUE,
                        include_health_conditions=TRUE,
                        include_language_exposure=TRUE,
                        include_study_internal_id=TRUE) %>%
                        filter(is_norming) -> participants
# More demographics data available, if needed -- modify select():
participants %>%
  select(child_id, age, sex) %>%
  write_csv(paste(filename_data, "participants.csv", sep="_"))
```

There should be two *csv* files:
`r paste(filename_data, "responses.csv", sep="_")` and
`r paste(filename_data, "participants.csv", sep="_")`.

The *responses* file is used to extract the list of `r scale` items
(with their definitions and English glosses, if available).
Relevant responses are recoded as 0's and 1's appropriately.

The *participants* file contains demographic data (*age* and *sex* in particular).

```{r data_import}
# Read csv files:
read_csv(paste(filename_data, "participants.csv", sep="_")) -> participants
read_csv(paste(filename_data, "responses.csv", sep="_")) -> responses

# Extract instrument contents (item definitions, English glosses, and IDs):
responses %>% distinct(item_definition, english_gloss, item_id) -> cdi

# Filter & recode responses based on scale:
if(subscale == "Gest") {
  responses %>% filter(str_starts(item_kind, "gestures")) %>%
      mutate(value = case_match(value,
                                c("not yet", "no") ~ 0,
                                c("often", "sometimes", "yes") ~ 1,
                                .default = NA)) -> responses
} else {
  if(subscale == "Prod") {
    positive <- c("produces", "understands")
  } else {
    positive <- "understands"
  }
  responses <- filter(responses, item_kind == "word") %>%
      mutate(value = case_match(value,
                                positive ~ 1,
                                NA ~ NA,
                                .default = 0)) -> responses
}

```

There are three key objects needed for all the analyses done and presented here:

1. `responses` matrix (administrations x items; 0/1)
2. `cdi` data frame (items with their definitions & glosses)
3. `scores` data frame (participant information: raw scores & demographics)

To prepare the `responses` matrix we eliminate *floor* and *ceiling* (zero variance) items.

**NB:** `Mirt` documentation suggests removing, as *floor* and *ceiling*,
items with difficulties > *.95* and < *.05*. **Check and explore in the future?**

```{r data_processing}
# Filter out floor & ceiling items:
original_n <- length(unique(responses$child_id))
responses %>% group_by(item_id) %>% filter(var(value) > 0) %>%
  pivot_wider(names_from = "item_id", values_from = "value", id_cols = "child_id") -> responses
# Remove them from the instrument data:
original_cdi_n <- nrow(cdi)
cdi %>% filter(item_id %in% colnames(responses)) -> cdi

# Calculate scores and merge with demographics:
responses %>% rowwise() %>% mutate(score = sum(across(starts_with("item_")))) %>%
     select(1, last_col()) %>% left_join(participants) %>% select(- last_col()) -> scores

# Create responses matrix:
responses <- as.matrix(select(responses, starts_with("item_")))
```

There are `r original_n` `r paste(language, scale)` administrations
and `r original_cdi_n` `r paste(language, scale)` items.
After removing *floor* and *ceiling* items there are
`r nrow(responses)` administrations and `r ncol(responses)` items
for IRT model fitting.


# Model creation

For the model creation
we start by deciding on the optimal number of quadrature points.
With a set number of quadrature points we proceed to fitting the model iteratively,
first removing any items without a fit,
and then removing all items showing high local dependence and fitting the model again.
We continue with removing misfits and locally dependent items if needed.

## Quadrature points

We decide on the number of quadrature points iteratively
by fitting models with increasing numbers and
comparing their fit. We start with the default and increase it in
a number of arbitrary, but hopefully not unreasonable, steps.

We also compare the parameter estimates for the first item,
just out of curosity: as the goodness of fit stabilises, so
should the estimates.

*It may take some time! Switch eval on only if you really need to recalculate.*

```{r quadpts_comp, eval=FALSE}
quadpts <- quadpts_comp(responses, quadpts = c(61, 151, 301, 451, 601, 751), NCYCLES = 100000,
             output_file = "Data/norwegian_ws_quadpts.RData")
```

We load the comparison of model fits for different numbers of quadrature points
from a file. Here there are:

```{r quadpts_comp1}
load("Data/norwegian_ws_quadpts.RData")
quadpts[1:3,]
```

And here are the parameter estimates for the first item:

```{r quadpts_comp2}
sapply(1:ncol(quadpts), function(x) quadpts[,x]$item1pars)
```

```{r quadpts_set}
quadtp <- 301
```

Based on the above we set the number of quadrature points to `r quadtp`.

## Model fitting, round 1

We iteratively fit the model and remove any misfits until we have fits for all items.

*It may take some time!*

We save the model to a file to save time in the future
(if you want to refit the model, switch the eval on):

```{r misfits_removal, eval=FALSE}
model <- misfits_removal(responses=responses, quadtps=quadtp, NCYCLES=100000, p=0.001,
                          output_file = "Data/norwegian_ws.RData")
```

We load the model from a file and print a basic summary:

```{r model_summary}
load("Data/norwegian_ws.RData")
mod <- model[[1]]
items_removed <- model[[2]]
rm(model) # to save space

mod
```

### Items fit

`r length(unlist(items_removed))` items were misfits in the process of model fitting and
were removed in `r length(items_removed)` iterations before a model with no misfits was obtained.

```{r items_removed_list}
cdi %>% filter(item_id %in% unlist(items_removed))
```

```{r items_removed}
if(length(items_removed)) {
  responses <- responses[, -which((colnames(responses) %in% unlist(items_removed)))]
  cdi %>% filter(! item_id %in% unlist(items_removed)) -> cdi
  if (nrow(cdi) == ncol(responses)) print(paste0(ncol(responses), " items left after removal."))
}
```


### Local dependence

*Check and explore the following code in the future (also compare with Kachergis et al., 2022)*

```{r local_dependence, eval=FALSE}
# I can't make the following code hide the results...
residuals <- residuals(mod, df.p=FALSE) # TODO: explore other parameters?
save(residuals, file = "Data/norwegian_ws_residuals.RData")
```

```{r local_dependence_histogram}
load("Data/norwegian_ws_residuals.RData")
residuals_up <- residuals
residuals_up[lower.tri(residuals_up)] <- NA # upper diagonal contains signed Cramers V coefficients
hist(residuals_up, main="Histogram of signed Cramer's V coefficients", xlab= "Cramer's V")
```

```{r local_dependence_summary}
summary(as.vector(residuals_up))[-7]
# -7 to avoid reporting NAs, which would be confusing as lower triangle is filled with them
```

To display the most highly mutually dependent items
we use V > .3 as the threshold:

```{r local_dependence_high}
indexes <- arrayInd(which(residuals_up > 0.3), dim(residuals))
data.frame(item1 = pull(cdi[indexes[, 1], "english_gloss"]),
           item2 = pull(cdi[indexes[, 2], "english_gloss"]), CramersV = residuals_up[indexes])
```

There are a few but following Kachergis et al. we set the threshold for removing items to V > .5
and we remove the following items:

```{r local_dependence_remove}
indexes <- arrayInd(which(residuals_up > 0.5), dim(residuals))
data.frame(item1 = pull(cdi[indexes[, 1], "item_definition"]),
           item2 = pull(cdi[indexes[, 2], "item_definition"]), CramersV = residuals_up[indexes])
unique(as.vector(indexes)) -> items_to_remove_ids
if(length(items_to_remove_ids)) {
     responses[, -items_to_remove_ids] -> responses
     cdi[-items_to_remove_ids,] -> cdi
}
```


## Model fitting, round 2

Then we repeat the iterative model fit with misfits removal.

```{r misfits_removal_2, eval=FALSE}
model <- misfits_removal(responses=responses, quadtps=quadtp, NCYCLES=100000, p=0.001,
                          output_file = "Data/norwegian_ws_2.RData")
```

We load the model from a file (comment out if just fitted above)
and print a basic summary:

```{r model_summary_2}
load("Data/norwegian_ws_2.RData")
mod <- model[[1]]
items_removed <- model[[2]]
rm(model) # to save space

mod
```


### Items fit

`r length(unlist(items_removed))` items were misfits in the process of model fitting and
were removed in `r length(items_removed)` iterations before a model with no misfits was obtained.

```{r items_removed_list_2}
cdi %>% filter(item_id %in% unlist(items_removed))
```

```{r items_removed_2}
if(length(items_removed)) {
  responses <- responses[, -which((colnames(responses) %in% unlist(items_removed)))]
  cdi %>% filter(! item_id %in% unlist(items_removed)) -> cdi
  if (nrow(cdi) == ncol(responses)) print(paste0(ncol(responses), " items left after removal."))
}
```


### Local dependence

*Seems like we don't need to repeat the local dependence procedure.*
*Cramer's Vs remain (almost)

```{r local_dependence_2, eval=FALSE}
# I can't make the following code hide the results...
residuals <- residuals(mod, df.p=FALSE) # TODO: explore other parameters?
save(residuals, file = "Data/norwegian_ws_residuals_2.RData")
```

```{r local_dependence_histogram_2, eval=FALSE}
load("Data/norwegian_ws_residuals_2.RData")
residuals_up <- residuals
residuals_up[lower.tri(residuals_up)] <- NA # upper diagonal contains signed Cramers V coefficients
hist(residuals_up, main="Histogram of signed Cramer's V coefficients", xlab= "Cramer's V")
```

```{r local_dependence_summary_2, eval=FALSE}
summary(as.vector(residuals_up))[-7]
# -7 to avoid reporting NAs, which would be confusing as lower triangle is filled with them
```

```{r local_dependence_high_2, eval=FALSE}
indexes <- arrayInd(which(residuals_up > 0.3), dim(residuals))
data.frame(item1 = pull(cdi[indexes[, 1], "english_gloss"]),
           item2 = pull(cdi[indexes[, 2], "english_gloss"]), CramersV = residuals_up[indexes])
```

```{r local_dependence_remove_2, eval=FALSE}
indexes <- arrayInd(which(residuals_up > 0.5), dim(residuals))
data.frame(item1 = pull(cdi[indexes[, 1], "item_definition"]),
           item2 = pull(cdi[indexes[, 2], "item_definition"]), CramersV = residuals_up[indexes])
unique(as.vector(indexes)) -> items_to_remove_ids
if(length(items_to_remove_ids)) {
     responses[, -items_to_remove_ids] -> responses
     cdi[-items_to_remove_ids,] -> cdi
}
```


# Model exploration

## IRT items parameters

```{r item_parameters}
params_IRT <- as.data.frame(coef(mod, simplify = TRUE, IRTpars = TRUE)$items)[1:2]
cdi <- cbind(params_IRT, cdi)
```

Scatterplot of difficulty by discrimination
(*we use English glosses;
also, if we had categories we could add them as colours*):

```{r item_parameters_plot, fig.width=12, fig.height=8}
# ggplot(params_cdi, aes(b, a, colour=category, label=position)) +
ggplot(cdi, aes(b, a, label=english_gloss)) +
  geom_text(alpha = 0.7, size=3) +
#  geom_point(alpha = 0.3) +
  xlim(-5, 3) +
  ylim(.5, 5.5) +
  xlab("Difficulty") +
  ylab("Discrimination") +
#  labs(colour = "Category") +
#  labs(title = "Norwegian CDI:WG Comprehension") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
  legend.position = "top"
)
```

## Ability vs. raw scores

We estimate ability levels for children from the sample based on the model.

```{r ability}
scores <- cbind(scores, fscores(mod, method = "MAP", full.scores.SE = TRUE))
```

*Explore the failed convergence? Does it have something to do with
the chosen MAP method?*

Plot the estimated ability level against raw scores:

```{r ability_scores_plot, fig.width=12, fig.height=8}
ggplot(scores, aes(score, F1)) +
  geom_point() +
  ylab("Ability level (theta)") +
  xlab("Raw score (sum of checked items)") +
  ylim(-4, 4) +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
)
```

The scatterplot above shows that IRT estimates get "normalised",
compared to raw scores, as can be seen on the following histograms
(flat distribution for raw scores vs. bell-like distribution for *theta* estimates).

```{r, eval=FALSE}
# Should be in two separate chunks perhaps?
hist(scores$score, xlab = "Raw sums", main = "")
hist(scores$F1, xlab = "Estimated ability levels", main = "")
```


## Ability vs. SE

Plot the ability level against SE:

```{r ability_se_plot, fig.width=12, fig.height=8}
ggplot(scores, aes(F1, SE_F1)) +
  geom_point(alpha = 0.3) +
  xlab("Ability level (theta)") +
  ylab("Standard error (SE)") +
  labs(title = "Norwegian CDI:WG Comprehension") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
)
```

The scatterplot above shows increasing errors at both extremes of the ability level.
Perhaps due to floor and ceiling effects. Errors higher than 0.1 highly unfrequent though:

```{r se_histogram, fig.width=12, fig.height=8}
hist(scores$SE_F1, main = "Norwegian CDI:WS", sub = "Histogram of SE", xlab = "SE")
```

We can export the data for crosslinguistic plotting/analysing.

```{r ability_se_export, eval=FALSE}
write_csv(scores, "Data/scores_Norwegian-WS.csv")
```


## SE and ability level vs. age

The graph below shows that higher error rates are at and close to age range limits
(which is sort of obvious, given the relation between theta and age).
But it also shows that some outliers are at every age
(though in the middle they are not as high).

```{r se_age_plot, fig.width=12, fig.height=8}
ggplot(data.frame(age = as.factor(scores$age), SE = scores$SE_F1), aes(age, SE)) +
  geom_boxplot() +
  xlab("Age (months)") +
  ylab("Standard error of ability estimations") +
  labs(title = "Norwegian CDI:WS") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
  )
```

So let's see the relation between ability level and age (and first between raw scores and age).
We add SE as colour.

```{r score_age_plot, fig.width=12, fig.height=8}
ggplot(scores) +
  geom_point(aes(age, score))
ggplot(scores) +
  geom_point(aes(age, F1, col = SE_F1)) +
  scale_fill_gradient(low="green", high="red", aesthetics = "col") +
  labs(y = "ability level", col = "SE") +
  ylim(-4, 4)
```


# Simulations

```{r prepare_sim}
params <- as.data.frame(coef(mod, simplify = T)$items)[1:2]
mo <- generate.mirt_object(params, '2PL')
cl <- makeCluster(detectCores())
```

## Whole pool for everyone

```{r sim_all, eval=FALSE}
message(paste("Starting simulation @ ", Sys.time()))
results_all <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI",
                       local_pattern = responses, cl = cl, design = list(min_items = ncol(responses)))
# results_all1 <- mirtCAT(mo = mod, method = "MAP", criteria = "MI", start_item = "MI",
#                       local_pattern = responses_r, cl = cl, design = list(min_items = ncol(responses_r)))
save(results_all, file = "Data/norwegian_ws_sim_all.Rdata")
message(paste("Finished @ ", Sys.time()))
beep()
```


```{r sim_all_report}
load("Data/norwegian_ws_sim_all.Rdata")
report_sim_results(results_all, select(scores, contains("F1")))
```

Based on the whole pool simulation we can examine decreasing average SE with each item
(we plot it up to 100 items: no need for more).

```{r se_length_plot, fig.width=12, fig.height=8, warning=FALSE}
SE_history <- laply(results_all, function(x) x$thetas_SE_history)
meanSE_item <- as.data.frame(colMeans(SE_history))
meanSE_item$item <- row.names(meanSE_item)
colnames(meanSE_item) <- c("mean_SE", "item")
meanSE_item$item <- as.numeric(meanSE_item$item)

ggline(meanSE_item, x = "item", y = "mean_SE", numeric.x.axis = TRUE) +
  xlab ("Test length") +
  ylab ("Mean SE") +
  labs(title = "Norwegian CDI:WS") +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "lightgrey"),
    text = element_text(size=16)
  ) + 
  scale_y_continuous(breaks = c(0.1, 0.15, 0.2, 0.25, 0.5, 0.75, 1)) + 
  scale_x_continuous(breaks = seq(0, nrow(params), by=10), limits = c(1, 100))
```

We can export the data for crosslinguistic plotting/analysing.

```{r mean_se_export, eval=FALSE}
write_csv(meanSE_item, "Data/meanSE_Norwegian-WS.csv")
```


## SE = 0.1 stop criterion

The plot above shows that we need 30-40 items to reach SE=.1 on average.
Let's try a simulation.

```{r se1, eval=FALSE}
results_SE1 <- sim_se(0.1, file = "Data/norwegian_ws_sim_se1.RData")
```


```{r se1_report}
load("Data/norwegian_ws_sim_se1.RData")
results_SE1 <- results
report_sim_results(results_SE1, select(scores, contains("F1")))
```


```{r se1_plot, fig.width=12, fig.height=8}
plot_length(results_SE1, "Norwegian CDI:WS", 0.1)
```

## SE = 0.15 stop criterion

```{r se15, eval=FALSE}
results_SE15 <- sim_se(0.15, file = "Data/norwegian_ws_sim_se15.RData")
```


```{r se15_report}
load("Data/norwegian_ws_sim_se15.RData")
results_SE15 <- results
report_sim_results(results_SE15, select(scores, contains("F1")))
```


```{r se15_plot, fig.width=12, fig.height=8}
plot_length(results_SE15, "Norwegian CDI:WS", 0.15)
```


## Different start thetas

```{r different_start_thetas, eval=FALSE, message=FALSE, echo=FALSE}
start_thetas <- get_start_thetas(scores, "age", "sex")
fscores_aggr <- start_thetas[[1]]
start_thetas <- start_thetas[[2]]
```

TODO: Simulations with different start thetas when stop criterion would be specified

# Input files preparation

```{r eval=FALSE}
prepare_input_files(params, cdi$item_definition, fscores_aggr,
                    question = "CATEGORY LABEL here", group = "comp", id = cdi$item_id,
                    file1 = "items-Norwegian-WS.csv", file2 = "startThetas-Norwegian-WS.csv")
```



