---
title: "Norwegian CDI:WG Gestures analysis"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE,
                      echo = FALSE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = "center",
                      fig.width = 6,
                      fig.height = 4)
```

# Setup

Below are the settings for the analyses reported in this document.

```{r settings, echo = TRUE}
language <- "Norwegian"
scale <- "WG Comprehension" # one of "WG Comprehension", "WG Production", "WG Gestures", "WS"
```

Plot titles, file names etc. are based on them.

```{r analyses_setup}
# Additional setup:
functions_dir <- "Functions_new" # "Functions_new" keeps updated function definition
data_dir <- "Data"

version <- strsplit(scale, " ")[[1]]
subscale <- "Prod"
if(length(version) > 1) substr(version[2], 1, 4) -> subscale
version <- version[1]
if(version == "WS") {
  filename_prefix <- tolower(paste(language, version, sep="_"))
} else {
  filename_prefix <- tolower(paste(language, version, subscale, sep="_"))
}
filename_prefix <- paste(data_dir, filename_prefix, sep="/")
filename_data <- paste(data_dir, tolower(paste(language, version, sep="_")), sep="/")
```

The following  libraries are loaded.

```{r libraries, echo=TRUE, results='hide'}
library(beepr)
library(tidyverse)
library(mirt)
library(mirtCAT)
library(parallel)
library(ggpubr)
```

Functions from "`r functions_dir`" directory are sourced.
Data are expected and saved in "`r data_dir`" directory.
 
```{r functions}
files <- list.files(path = functions_dir, pattern = "\\.R$", full.names = TRUE)
for (f in files) source(f)
```


## Data preparation

Data may be downloaded from the Wordbank database and saved locally
for quicker access in the future. See and run the chunk below for that
(modifying calls to `select()` if more data, demographics in particular, is needed).

```{r wordbank_download, eval=FALSE, echo=TRUE}
# Get responses and item data:
wordbankr::get_instrument_data(language, version,
                               administration_info=TRUE,
                               item_info=TRUE) %>%
     filter(is_norming) -> responses
# More item data available, if needed -- modify select():
responses %>%
     select(child_id, item_id, item_definition, english_gloss, item_kind, value) %>%
     write_csv(paste(filename_data, "responses.csv", sep="_"))

# Get demographics data:
wordbankr::get_administration_data(language, version,
                                   include_demographic_info=TRUE,
                                   include_birth_info=TRUE,
                                   include_health_conditions=TRUE,
                                   include_language_exposure=TRUE,
                                   include_study_internal_id=TRUE) %>%
     filter(is_norming) -> participants
# More demographics data available, if needed -- modify select():
participants %>%
     select(child_id, age, sex) %>%
     write_csv(paste(filename_data, "participants.csv", sep="_"))
```

There should be two *csv* files:

1. "`r paste(filename_data, "responses.csv", sep="_")`" and
2. "`r paste(filename_data, "participants.csv", sep="_")`".

First one is used to extract the list of `r paste(language, scale)` items
(with their definitions and English glosses, if available).
Relevant responses are recoded as 0's and 1's appropriately.

Second one contains demographic data (*age* and *sex* in particular).

```{r data_import}
# Read csv files:
read_csv(paste(filename_data, "participants.csv", sep="_")) -> participants
read_csv(paste(filename_data, "responses.csv", sep="_")) -> responses

# Filter & recode responses based on scale:
if(subscale == "Gest") {
     responses %>% filter(str_starts(item_kind, "gestures")) %>%
          mutate(value = case_match(value,
                                   c("not yet", "no") ~ 0,
                                   c("often", "sometimes", "yes") ~ 1,
                                   .default = 0)) -> responses
} else {
     if(subscale == "Prod") {
          positive <- c("produces", "understands")
     } else {
          positive <- "understands"
     }
     responses %>% filter(item_kind == "word") %>%
          mutate(value = case_match(value,
                                    positive ~ 1,
                                    NA ~ 0,
                                    .default = 0)) -> responses
}

# Extract instrument contents (item definitions, English glosses, and IDs):
responses %>% distinct(item_definition, english_gloss, item_id) -> cdi
```

There are three key objects needed for all the analyses done and presented here:

1. `responses` matrix (administrations x items; 0/1)
2. `cdi` data frame (items with their definitions & glosses)
3. `scores` data frame (participant information: raw scores & demographics)

To prepare the `responses` matrix we eliminate *floor* and *ceiling* (zero variance) items.

**NB:** `Mirt` documentation suggests removing, as *floor* and *ceiling*,
items with difficulties > *.95* and < *.05*. **Check and explore in the future?**

```{r data_processing}
# Filter out floor & ceiling items:
original_n <- length(unique(responses$child_id))
responses %>% group_by(item_id) %>% filter(var(value) > 0) %>%
     pivot_wider(names_from = "item_id", values_from = "value", id_cols = "child_id") -> responses
# Remove them from the instrument data:
original_cdi_n <- nrow(cdi)
cdi %>% filter(item_id %in% colnames(responses)) -> cdi

# Calculate scores and merge with demographics:
responses %>% rowwise() %>% mutate(score = sum(across(starts_with("item_")))) %>%
     select(1, last_col()) %>% left_join(participants) %>% select(- last_col()) -> scores

# Create responses matrix:
responses <- as.matrix(select(responses, starts_with("item_")))
```

There are `r original_n` `r paste(language, scale)` administrations
and `r original_cdi_n` `r paste(language, scale)` items.
After removing *floor* and *ceiling* items there are
`r nrow(responses)` administrations and `r ncol(responses)` items
for IRT model fitting.


# Model creation

For the model creation
we start by deciding on the optimal number of quadrature points.
With a set number of quadrature points we proceed to fitting the model iteratively,
first removing any items without a fit,
and then removing all items showing high local dependence and fitting the model again.
We continue with removing misfits and locally dependent items if needed.

## Quadrature points

### Introduction

The number of quadrature points (`quadpts` parameter in `mirt` package)
is a parameter important for numerical integration used
in the IRT model fitting. We know that for some of our datasets the default number
results in the model not converging. Therefore with new data we used to
routinely first try model fitting with an increasing number of `quadpts`,
starting with the default 61, and going up by an arbitrary number of arbitrary
steps. We would then compare the convergence of the model, its fit, and
the parameter estimates for the first item, and choose the `quadpts`
for which these statistics stabilised (roughly).

Given that model fitting is a one-off thing (we save the model for later reuse),
it seems more convenient -- and also beneficial in the context of cross data comparisons --
to set `quadpts` reasonably high and keep it constant. The following section
is totally optional.

### Model comparisons (only if needed)

From time to time we might want to run these comparisons. E.g., if new analyses
yield results different from the old ones, or for new data, especially if
the model doesn't converge on the first try.

The `quadpts` values tested and the choice criteria are not precise but
it shouldn't matter. The greater the value the longer fitting time but greater precision.
The default, historical, vector `quadpts = c(61, 151, 301, 451, 601, 751)` can be
changed as required.

```{r quadpts_comp, eval=FALSE, echo=TRUE}
quadpts <- quadpts_comp(responses, quadpts = c(61, 151, 301, 451, 601, 751), NCYCLES = 100000,
             output_file = paste(filename_prefix, "quadpts.RData", sep="_"))
```

Results of the above code are saved to a file
("`r paste(filename_prefix, "quadpts.RData", sep="_")`"), so it needs to be run
only once.

The convergence status and model fits (logLik) for different values of `quadpts`
(which can be printed below) are read from that file.

```{r quadpts_comp_fit, eval=FALSE}
load(paste(filename_prefix, "quadpts.RData", sep="_"))
quadpts[1:3,]
```

The same applies to parameter estimates for the first item.

```{r quadpts_comp_estimates, eval=FALSE}
sapply(1:ncol(quadpts), function(x) quadpts[,x]$item1pars)
```


### Setting

Based on the above (**or not**) we set the number of quadrature points.

```{r quadpts_set, echo=TRUE}
quadtp <- 600
```


## Model fitting (**iterate**, if needed)

When fitting the IRT model, we eliminate items for which:

1. the model doesn't fit (significant S-X²);
2. the residuals are too closely related to another item's residuals (Cramer's $V > .5$).

If point 2 applies this whole section has to be re**iterated**.
In such a case modify `fitting_iteration` accordingly for future *kniting*.

```{r fitting_iteration, echo=TRUE}
fitting_iteration <- 1
```

Some of the code chunks in this section may take time to run, so their
output is saved to a file, with the **iteration** number attached
to the filename. Rerun them only if needed.

### Misfits removal

First we repeatedly fit the model with `r quadtp` quadrature points,
each time remove any misfits until we have fits for all items.
The final model and the list of removed items are saved to
"`r paste(filename_prefix, fitting_iteration, "model.RData", sep="_")`" file.

```{r misfits_removal, eval=FALSE}
model <- misfits_removal(responses=responses, quadtps=quadtp, NCYCLES=100000, p=0.001,
                          output_file = paste(filename_prefix, fitting_iteration, "model.RData", sep="_"))
```

We print the model's basic summary.

```{r model_summary}
load(paste(filename_prefix, fitting_iteration, "model.RData", sep="_"))
mod <- model[[1]]
items_removed <- model[[2]]
item_fit <- model[[3]]
rm(model) # to save space

mod
```

`r length(unlist(items_removed))` items were misfits in the process of model fitting.

```{r items_removed_list}
cdi %>% filter(item_id %in% unlist(items_removed))
```

They were removed in `r length(items_removed)` iterations before a model with no misfits was obtained
(misfits removing function iterations, not to be confused with `fitting_iteration`
equaled `r fitting_iteration`).

```{r items_removed}
if(length(items_removed)) {
  responses <- responses[, -which((colnames(responses) %in% unlist(items_removed)))]
  cdi %>% filter(! item_id %in% unlist(items_removed)) -> cdi
  if (nrow(cdi) == ncol(responses)) print(paste0(ncol(responses), " items left after removal."))
}
```


### Local dependence

As the next step we check pairwise correlations between
items after taking the model (i.e., the underlying trait) into account.
For each pair of residuals we calculate Cramer's $V$, a measure of
association between two dichotomous variables.

#### More on the issue

What `mirt` package returns are **signed** coefficients and the sign indicates the direction of
the relation. Positive relation (*local dependence*) means that two items
are more related than the model predicts, i.e., are too similar or redundant.
IRT assumes *local independence*: after conditioning on the latent trait $\theta$,
responses are assumed to be statistically independent; or in plain English,
if we know a person’s ability level, knowing their response to one item
should not help us predict their response to another; *Chat GPT* assistance).

Keeping locally dependent items in the item pool may lead to a number of
distortions in CAT measurement (crucially, among other, inflated $\theta$ estimates;
*Chat GPT* assistance). This is why it is important to monitor this problem.
As with any correlation metric however, its interpretation is always arbitrary.

So far we have simply removed all item pairs with $V > .5$.
However, $.5$ might be too *high* as a threshold, potentially leaving in too many
locally dependent items. At the same time, removing *both items* is obviously
too harsh, reducing the item pool excessively,
since the problem is the relation between them, not the items themselves.

Deciding on which item to keep is not straightforward though.
The simplest option is to rely on a single statistic, e.g., *item's fit*,
but high $V$ pairs are often *gender pairs* like "mommy--daddy", "boy--girl",
"grandpa--granny", "his--her": keeping only one seems somewhat questionable
(this is in fact a potential guideline for future CDI development/revisions:
**make such pairs single items**).
Moreover, the same item may appear in a number of high $V$ pairs:
it's reasonable to remove that item, even if its fit is better than the other items'.

Finally, so far we have overlooked negative $V$'s:
they suggest the violation of the local independence assumption as well
(hidden multidimensionality might be at play).
However, they seem to be less common and might be less distortive for CAT measurement
(*Chat GPT* assistance; further checks needed).

All the above suggests that diagnosing local dependence is a complex issue that
might require careful examination and deliberate decisions, rether than
a purely algorithmic procedure.

#### Summary & selection

We calculate Cramer's $V$ for each pair of residuals and save the results to
"`r paste(filename_prefix, fitting_iteration, "residuals.RData", sep="_")`" file.

```{r local_dependence, eval=FALSE}
# I can't make the following code hide the results...
residuals <- residuals(mod, df.p=FALSE) # TODO: explore other parameters?
save(residuals, file = paste(filename_prefix, fitting_iteration, "residuals.RData", sep="_"))
```

The histogram below shows the (**signed**) Cramer's $V$ distribution.

```{r local_dependence_histogram}
load(paste(filename_prefix, fitting_iteration, "residuals.RData", sep="_"))
residuals_up <- residuals
residuals_up[lower.tri(residuals_up)] <- NA # upper diagonal contains signed Cramers V coefficients
hist(residuals_up, main="Histogram of signed Cramer's V coefficients", xlab= "Cramer's V")
```

These are the descriptive statistics of the distribution.

```{r local_dependence_summary}
summary(as.vector(residuals_up))[-7]
# -7 to avoid reporting NAs, which would be confusing as lower triangle is filled with them
```

To display *questionably* mutually dependent items
we use $V > .3$ as the threshold. We print their definitions, English glosses,
and fit statistics (\mathrm{S\!-\!X}^2).

```{r local_dependence_positive}
cramers_V_df(arrayInd(which(residuals_up > 0.3), dim(residuals)), cdi, item_fit) %>%
  select(- starts_with("item_id"))
```

In the same way we also print all pairs with *questionably* high negative
values ($V < -.3$).

```{r local_dependence_negative}
cramers_V_df(arrayInd(which(residuals_up < -0.3), dim(residuals)), cdi, item_fit) %>%
  select(- starts_with("item_id"))
```


To **remove** the most highly mutually dependent items
we use $V > .5$ as the threshold.

```{r local_dependence_remove}
indexes <- arrayInd(which(residuals_up > 0.5), dim(residuals))
cramers_V_df(indexes, cdi, item_fit) %>%
  select(- starts_with("item_id"))
unique(as.vector(indexes)) -> items_to_remove_ids
if(length(items_to_remove_ids)) {
     responses[, -items_to_remove_ids] -> responses
     cdi[-items_to_remove_ids,] -> cdi
}
```

Number of removed items is `r length(items_to_remove_ids)`.
If it's greater than zero we need to increase the `fitting_iteration` number
by one (see chunk below) and repeat everything from
the [Misfits removal](#misfits-removal) section onwards
(also don't forget to update the `fitting_iteration` chunk right before
[Misfits removal](#misfits-removal)).

```{r fitting_iteration_increment, eval=FALSE, echo=TRUE}
fitting_iteration <- fitting_iteration + 1
```


# Model exploration

## IRT items parameters

```{r item_parameters}
params_IRT <- as.data.frame(coef(mod, simplify = TRUE, IRTpars = TRUE)$items)[1:2]
cdi <- cbind(params_IRT, cdi)
```

Scatterplot of difficulty by discrimination
(*we use English glosses;
also, if we had categories we could add them as colours*):

```{r item_parameters_plot, fig.width=12, fig.height=8}
# ggplot(params_cdi, aes(b, a, colour=category, label=position)) +
ggplot(cdi, aes(b, a, label=english_gloss)) +
  geom_text(alpha = 0.7, size=3) +
#  geom_point(alpha = 0.3) +
  xlim(-5, 3) +
  ylim(.5, 5.5) +
  xlab("Difficulty") +
  ylab("Discrimination") +
#  labs(colour = "Category") +
#  labs(title = "Norwegian CDI:WG Comprehension") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
  legend.position = "top"
)
```

## Ability vs. raw scores

We estimate ability levels for children from the sample based on the model.

```{r ability}
scores <- cbind(scores, fscores(mod, method = "MAP", full.scores.SE = TRUE))
```

*Explore the failed convergence? Does it have something to do with
the chosen MAP method?*

Plot the estimated ability level against raw scores:

```{r ability_scores_plot, fig.width=12, fig.height=8}
ggplot(scores, aes(score, F1)) +
  geom_point() +
  ylab("Ability level (theta)") +
  xlab("Raw score (sum of checked items)") +
  ylim(-4, 4) +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
)
```

The scatterplot above shows that IRT estimates get "normalised",
compared to raw scores, as can be seen on the following histograms
(flat distribution for raw scores vs. bell-like distribution for *theta* estimates).

```{r, eval=FALSE}
# Should be in two separate chunks perhaps?
hist(scores$score, xlab = "Raw sums", main = "")
hist(scores$F1, xlab = "Estimated ability levels", main = "")
```


## Ability vs. SE

Plot the ability level against SE:

```{r ability_se_plot, fig.width=12, fig.height=8}
ggplot(scores, aes(F1, SE_F1)) +
  geom_point(alpha = 0.3) +
  xlab("Ability level (theta)") +
  ylab("Standard error (SE)") +
  labs(title = "Norwegian CDI:WG Comprehension") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
)
```

The scatterplot above shows increasing errors at both extremes of the ability level.
Perhaps due to floor and ceiling effects. Errors higher than 0.1 highly unfrequent though:

```{r se_histogram, fig.width=12, fig.height=8}
hist(scores$SE_F1, main = "Norwegian CDI:WS", sub = "Histogram of SE", xlab = "SE")
```

We can export the data for crosslinguistic plotting/analysing.

```{r ability_se_export, eval=FALSE}
write_csv(scores, "Data/scores_Norwegian-WS.csv")
```


## SE and ability level vs. age

The graph below shows that higher error rates are at and close to age range limits
(which is sort of obvious, given the relation between theta and age).
But it also shows that some outliers are at every age
(though in the middle they are not as high).

```{r se_age_plot, fig.width=12, fig.height=8}
ggplot(data.frame(age = as.factor(scores$age), SE = scores$SE_F1), aes(age, SE)) +
  geom_boxplot() +
  xlab("Age (months)") +
  ylab("Standard error of ability estimations") +
  labs(title = "Norwegian CDI:WS") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "lightgrey"),
  text = element_text(size=16)
  )
```

So let's see the relation between ability level and age (and first between raw scores and age).
We add SE as colour.

```{r score_age_plot, fig.width=12, fig.height=8}
ggplot(scores) +
  geom_point(aes(age, score))
ggplot(scores) +
  geom_point(aes(age, F1, col = SE_F1)) +
  scale_fill_gradient(low="green", high="red", aesthetics = "col") +
  labs(y = "ability level", col = "SE") +
  ylim(-4, 4)
```


# Simulations

```{r prepare_sim}
params <- as.data.frame(coef(mod, simplify = T)$items)[1:2]
mo <- generate.mirt_object(params, '2PL')
cl <- makeCluster(detectCores())
```

## Whole pool for everyone

```{r sim_all, eval=FALSE}
message(paste("Starting simulation @ ", Sys.time()))
results_all <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI",
                       local_pattern = responses, cl = cl, design = list(min_items = ncol(responses)))
# results_all1 <- mirtCAT(mo = mod, method = "MAP", criteria = "MI", start_item = "MI",
#                       local_pattern = responses_r, cl = cl, design = list(min_items = ncol(responses_r)))
save(results_all, file = "Data/norwegian_ws_sim_all.Rdata")
message(paste("Finished @ ", Sys.time()))
beep()
```


```{r sim_all_report}
load("Data/norwegian_ws_sim_all.Rdata")
report_sim_results(results_all, select(scores, contains("F1")))
```

Based on the whole pool simulation we can examine decreasing average SE with each item
(we plot it up to 100 items: no need for more).

```{r se_length_plot, fig.width=12, fig.height=8, warning=FALSE}
SE_history <- laply(results_all, function(x) x$thetas_SE_history)
meanSE_item <- as.data.frame(colMeans(SE_history))
meanSE_item$item <- row.names(meanSE_item)
colnames(meanSE_item) <- c("mean_SE", "item")
meanSE_item$item <- as.numeric(meanSE_item$item)

ggline(meanSE_item, x = "item", y = "mean_SE", numeric.x.axis = TRUE) +
  xlab ("Test length") +
  ylab ("Mean SE") +
  labs(title = "Norwegian CDI:WS") +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "lightgrey"),
    text = element_text(size=16)
  ) + 
  scale_y_continuous(breaks = c(0.1, 0.15, 0.2, 0.25, 0.5, 0.75, 1)) + 
  scale_x_continuous(breaks = seq(0, nrow(params), by=10), limits = c(1, 100))
```

We can export the data for crosslinguistic plotting/analysing.

```{r mean_se_export, eval=FALSE}
write_csv(meanSE_item, "Data/meanSE_Norwegian-WS.csv")
```


## SE = 0.1 stop criterion

The plot above shows that we need 30-40 items to reach SE=.1 on average.
Let's try a simulation.

```{r se1, eval=FALSE}
results_SE1 <- sim_se(0.1, file = "Data/norwegian_ws_sim_se1.RData")
```


```{r se1_report}
load("Data/norwegian_ws_sim_se1.RData")
results_SE1 <- results
report_sim_results(results_SE1, select(scores, contains("F1")))
```


```{r se1_plot, fig.width=12, fig.height=8}
plot_length(results_SE1, "Norwegian CDI:WS", 0.1)
```

## SE = 0.15 stop criterion

```{r se15, eval=FALSE}
results_SE15 <- sim_se(0.15, file = "Data/norwegian_ws_sim_se15.RData")
```


```{r se15_report}
load("Data/norwegian_ws_sim_se15.RData")
results_SE15 <- results
report_sim_results(results_SE15, select(scores, contains("F1")))
```


```{r se15_plot, fig.width=12, fig.height=8}
plot_length(results_SE15, "Norwegian CDI:WS", 0.15)
```


## Different start thetas

```{r different_start_thetas, eval=FALSE, message=FALSE, echo=FALSE}
start_thetas <- get_start_thetas(scores, "age", "sex")
fscores_aggr <- start_thetas[[1]]
start_thetas <- start_thetas[[2]]
```

TODO: Simulations with different start thetas when stop criterion would be specified

# Input files preparation

```{r eval=FALSE}
prepare_input_files(params, cdi$item_definition, fscores_aggr,
                    question = "CATEGORY LABEL here", group = "comp", id = cdi$item_id,
                    file1 = "items-Norwegian-WS.csv", file2 = "startThetas-Norwegian-WS.csv")
```



