---
title: "CAT simulations"
subtitle: "Norwegian CDI:WG Comprehension"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    code_folding: show
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE,
                      echo = FALSE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = "center",
                      fig.width = 6,
                      fig.height = 4)
```

# Setup

Below are the settings for the analyses reported in this document.

```{r settings, echo = TRUE}
language <- "Norwegian"
scale <- "WG Comprehension" # one of "WG Comprehension", "WG Production", "WG Gestures", "WS"
```

Plot titles, file names etc. are based on them.

```{r analyses_setup}
# Additional setup:
functions_dir <- "Functions_new" # "Functions_new" keeps updated function definition
data_dir <- "Data"

version <- strsplit(scale, " ")[[1]]
subscale <- "Prod"
if(length(version) > 1) substr(version[2], 1, 4) -> subscale
version <- version[1]
if(version == "WS") {
  filename_prefix <- tolower(paste(language, version, sep="_"))
} else {
  filename_prefix <- tolower(paste(language, version, subscale, sep="_"))
}
filename_prefix <- paste(data_dir, filename_prefix, sep="/")
```

The following libraries are loaded.

```{r libraries, echo=TRUE, results='hide'}
library(beepr)
library(tidyverse)
library(mirtCAT)
library(parallel)
#library(ggpubr)
```

Functions from "`r functions_dir`" directory are sourced.
Data are expected and saved in "`r data_dir`" directory.
 
```{r functions}
files <- list.files(path = functions_dir, pattern = "\\.R$", full.names = TRUE)
for (f in files) source(f)
```


## Data loading

There are four key objects needed for all the analyses done and presented here:

1. IRT model (`mod`)
1. `responses` matrix (administrations x items; 0/1), used for the final model (`mod`) fitting
2. `cdi` data frame (items with their definitions & glosses), used for the final model (`mod`) fitting
3. `scores` data frame (participant information: raw scores & demographics)


```{r data_loading}
list.files(path = data_dir, pattern = "data\\.RData$", full.names = TRUE) |>
  str_match(paste0(filename_prefix, "_([0-9]+)_")) |>
  as.numeric() |> na.omit() |> max() -> last_iteration

scores <- read_csv(paste(filename_prefix, "scores.csv", sep="_"))

load(paste(filename_prefix, last_iteration, "data.RData", sep="_"))
load(paste(filename_prefix, last_iteration, "model.RData", sep="_"))
  
mod <- model[[1]]
rm(model) # to save space
```


# Model exploration

## IRT items parameters

```{r item_parameters}
params_IRT <- as.data.frame(coef(mod, simplify = TRUE, IRTpars = TRUE)$items)[1:2]
cdi <- cbind(params_IRT, cdi)
```

We use a scatterplot to plot item parameters
(*we use English glosses;
also, if we had categories we could add them as colours*).

```{r item_parameters_plot}
# ggplot(params_cdi, aes(b, a, colour=category, label=position)) +
ggplot(cdi, aes(b, a, label=english_gloss)) +
  geom_text(alpha = 0.7, size=3) +
#  geom_point(alpha = 0.3) +
#  xlim(-5, 3) +
#  ylim(.5, 5.5) +
  xlab("Difficulty") +
  ylab("Discrimination") +
#  labs(colour = "Category") +
  labs(title = paste(language, scale)) +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "grey"),
    legend.position = "top"
  )
```

## Ability vs raw scores

We compute ability estimates ($\theta$'s) for children from the sample.

**NB:** For some datasets (models) the computation does not converge: should be looked into
(does it have something to do with the chosen MAP method?).

```{r ability}
scores <- cbind(scores, fscores(mod, method = "MAP", full.scores.SE = TRUE))
```

We plot estimated abilities ($\theta$'s) against raw scores (sums of checked items).

```{r ability_scores_plot}
ggplot(scores, aes(score, F1)) +
  geom_point() +
  ylab(expression(theta)) +
  xlab("Raw score") +
  labs(title = paste(language, scale)) +
#  ylim(-4, 4) +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "lightgrey"),
    text = element_text(size=16)
  )
```

The scatterplot above shows that IRT estimates get "normalised",
compared to raw scores.
The same can be seen comparing the following histograms.

```{r ability_scores_histograms}
# Should be in two separate chunks perhaps?
hist(scores$score, xlab = "Raw score", main = paste(language, scale))
hist(scores$F1, xlab = expression(theta), main = paste(language, scale))
```


## Ability vs SE

We plot ability estimates against their standard errors.

```{r ability_se_plot, fig.width=12, fig.height=8}
ggplot(scores, aes(F1, SE_F1)) +
  geom_point(alpha = 0.3) +
  xlab(expression(theta)) +
  ylab(expression(SE[theta])) +
  labs(title = paste(language, scale)) +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "lightgrey"),
    text = element_text(size=16),
    axis.title = element_text(size=20)
  )
```

The scatterplot above shows increasing errors at both extremes of the ability level
(perhaps due to floor and ceiling effects).
We also plot the distribution of SEs.

```{r se_histogram, fig.width=12, fig.height=8}
hist(scores$SE_F1, main = paste(language, scale), xlab = expression(SE[theta]))
```

We can export the data for crosslinguistic plotting/analysing.

```{r ability_se_export, eval=FALSE}
write_csv(scores, paste(filename_prefix, "thetas.csv", sep="_"))
```


## SE and ability level vs age

The graph below shows distribution of standard errors by age.

```{r se_age_plot}
ggplot(data.frame(age = as.factor(scores$age), SE = scores$SE_F1), aes(age, SE)) +
  geom_boxplot() +
  xlab("Age (months)") +
  ylab(expression(SE[theta])) +
  labs(title = paste(language, scale)) +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "lightgrey"),
    text = element_text(size=16)
  )
```

Now we plot the distribution of raw scores by age...

```{r score_age_plot}
ggplot(scores) +
  geom_point(aes(age, score)) +
  xlab("Age (months)") +
  ylab("Raw score") +
  labs(title = paste(language, scale))
```

...and the distribution of ability by age with colour indicating the ability's SE.

```{r ability_age_plot}
ggplot(scores) +
  geom_point(aes(age, F1, col = SE_F1)) +
  scale_fill_gradient(low="green", high="red", aesthetics = "col") +
  xlab("Age (months)") +
  ylab(expression(theta)) +
  labs(title = paste(language, scale), col = expression(SE[theta]))
#  ylim(-4, 4)
```


# Simulations

```{r prepare_sim}
params <- as.data.frame(coef(mod, simplify = T)$items)[1:2]
mo <- generate.mirt_object(params, '2PL')
cl <- makeCluster(detectCores())
```

## Whole pool for everyone

```{r sim_all, eval=FALSE}
message(paste("Starting simulation @ ", Sys.time()))
results_all <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI",
                       local_pattern = responses, cl = cl, design = list(min_items = ncol(responses)))
# results_all1 <- mirtCAT(mo = mod, method = "MAP", criteria = "MI", start_item = "MI",
#                       local_pattern = responses_r, cl = cl, design = list(min_items = ncol(responses_r)))
save(results_all, file = "Data/norwegian_ws_sim_all.Rdata")
message(paste("Finished @ ", Sys.time()))
beep()
```


```{r sim_all_report}
load("Data/norwegian_ws_sim_all.Rdata")
report_sim_results(results_all, select(scores, contains("F1")))
```

Based on the whole pool simulation we can examine decreasing average SE with each item
(we plot it up to 100 items: no need for more).

```{r se_length_plot, fig.width=12, fig.height=8, warning=FALSE}
SE_history <- laply(results_all, function(x) x$thetas_SE_history)
meanSE_item <- as.data.frame(colMeans(SE_history))
meanSE_item$item <- row.names(meanSE_item)
colnames(meanSE_item) <- c("mean_SE", "item")
meanSE_item$item <- as.numeric(meanSE_item$item)

ggline(meanSE_item, x = "item", y = "mean_SE", numeric.x.axis = TRUE) +
  xlab ("Test length") +
  ylab ("Mean SE") +
  labs(title = "Norwegian CDI:WS") +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "lightgrey"),
    text = element_text(size=16)
  ) + 
  scale_y_continuous(breaks = c(0.1, 0.15, 0.2, 0.25, 0.5, 0.75, 1)) + 
  scale_x_continuous(breaks = seq(0, nrow(params), by=10), limits = c(1, 100))
```

We can export the data for crosslinguistic plotting/analysing.

```{r mean_se_export, eval=FALSE}
write_csv(meanSE_item, "Data/meanSE_Norwegian-WS.csv")
```


## SE = 0.1 stop criterion

The plot above shows that we need 30-40 items to reach SE=.1 on average.
Let's try a simulation.

```{r se1, eval=FALSE}
results_SE1 <- sim_se(0.1, file = "Data/norwegian_ws_sim_se1.RData")
```


```{r se1_report}
load("Data/norwegian_ws_sim_se1.RData")
results_SE1 <- results
report_sim_results(results_SE1, select(scores, contains("F1")))
```


```{r se1_plot, fig.width=12, fig.height=8}
plot_length(results_SE1, "Norwegian CDI:WS", 0.1)
```

## SE = 0.15 stop criterion

```{r se15, eval=FALSE}
results_SE15 <- sim_se(0.15, file = "Data/norwegian_ws_sim_se15.RData")
```


```{r se15_report}
load("Data/norwegian_ws_sim_se15.RData")
results_SE15 <- results
report_sim_results(results_SE15, select(scores, contains("F1")))
```


```{r se15_plot, fig.width=12, fig.height=8}
plot_length(results_SE15, "Norwegian CDI:WS", 0.15)
```


## Different start thetas

```{r different_start_thetas, eval=FALSE, message=FALSE, echo=FALSE}
start_thetas <- get_start_thetas(scores, "age", "sex")
fscores_aggr <- start_thetas[[1]]
start_thetas <- start_thetas[[2]]
```

TODO: Simulations with different start thetas when stop criterion would be specified

# Input files preparation

```{r eval=FALSE}
prepare_input_files(params, cdi$item_definition, fscores_aggr,
                    question = "CATEGORY LABEL here", group = "comp", id = cdi$item_id,
                    file1 = "items-Norwegian-WS.csv", file2 = "startThetas-Norwegian-WS.csv")
```



