---
title: "CAT simulations"
subtitle: "`r paste(params$language, params$scale)`"
params:
  knitting:
    label: "You shouldn't knit directly from RStudio. Use the 'knit_from_template.R' file. Check only if you really want to override."
    value: FALSE
  language:
    label: "Choose language (preferably as appears in Wordbank):"
    value: ""
  scale:
    label: "Choose scale:"
    value: NULL
    choices: ["WG Comprehension", "WG Production", "WG Gestures", "WS"]
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    code_folding: show
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE,
                      echo = FALSE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = "center",
                      fig.width = 12,
                      fig.height = 8)
if(! params$knitting & isTRUE(getOption("knitr.in.progress"))) {
  stop("Don't knit directly from RStudio. Use the 'knit_from_template.R' file.")
}
if(params$language == "" | is.null(params$scale)) {
  # Ask for params
  pick_params <- function() {
    ui <- miniUI::miniPage(
      miniUI::gadgetTitleBar("Dataset details"),
      miniUI::miniContentPanel(
        shiny::textInput("language", "Language (preferably as appears in Wordbank):", value = ""),
        shiny::radioButtons("scale", "Scale:",
                            choices = c("WG Comprehension", "WG Production", "WG Gestures", "WS"))
      )
    )
    server <- function(input, output, session) {
      shiny::observeEvent(input$done, {
        shiny::stopApp(list(language = input$language, scale = input$scale))
      })
    }
    shiny::runGadget(ui, server)
  }
  params <- pick_params()
}
language <- params$language
scale    <- params$scale
```

# Setup

Do not change the code or text in the document, unless you know what you are doing or
you encounter an explicit encouragement to tweak code or settings.
No modifications should be required to run the code (and eventually knit the document),
as long as you successfully run the IRT model fitting for this dataset first.

```{r analyses_setup}
# Additional setup:
functions_dir <- "Functions"
data_dir <- "Data"

version <- strsplit(scale, " ")[[1]]
subscale <- "Prod"
if(length(version) > 1) substr(version[2], 1, 4) -> subscale
version <- version[1]
if(version == "WS") {
  filename_prefix <- tolower(paste(language, version, sep="_"))
} else {
  filename_prefix <- tolower(paste(language, version, subscale, sep="_"))
}
filename_prefix <- paste(data_dir, filename_prefix, sep="/")
```

The following libraries are loaded.

```{r libraries, echo=TRUE, results='hide'}
library(beepr)
library(tidyverse)
library(mirtCAT)
library(parallel)
```

Functions from "`r functions_dir`" directory are sourced.
Data are expected and saved in "`r data_dir`" directory.
 
```{r functions}
files <- list.files(path = functions_dir, pattern = "\\.R$", full.names = TRUE)
for (f in files) source(f)
```


## Data loading

There are four key objects needed for all the analyses done and presented here:

1. IRT model (`mod`)
1. `responses` matrix (administrations x items; 0/1), used for the final model (`mod`) fitting
2. `cdi` data frame (items with their definitions & glosses), used for the final model (`mod`) fitting
3. `scores` data frame (participant information: raw scores & demographics)


```{r data_loading}
list.files(path = data_dir, pattern = "data\\.RData$", full.names = TRUE) |>
  str_match(paste0(filename_prefix, "_([0-9]+)_")) |>
  as.numeric() |> na.omit() |> max() -> last_iteration

scores <- read_csv(paste(filename_prefix, "scores.csv", sep="_"))

load(paste(filename_prefix, last_iteration, "data.RData", sep="_"))
load(paste(filename_prefix, last_iteration, "model.RData", sep="_"))
  
mod <- model[[1]]
rm(model) # to save space
```


# Model exploration

The model was fit to `r dim(mod@Data$data)[2]` items
and `r dim(mod@Data$data)[1]` administrations.

## IRT items parameters

```{r item_parameters}
params_IRT <- as.data.frame(coef(mod, simplify = TRUE, IRTpars = TRUE)$items)[1:2]
cdi <- cbind(params_IRT, cdi)
```

We use a scatterplot to plot item parameters
(*we use English glosses;
also, if we had categories we could add them as colours*).

```{r item_parameters_plot}
# ggplot(params_cdi, aes(b, a, colour=category, label=position)) +
ggplot(cdi, aes(b, a, label=english_gloss)) +
  geom_text(alpha = 0.7, size=3) +
#  geom_point(alpha = 0.3) +
#  xlim(-5, 3) +
#  ylim(.5, 5.5) +
  xlab("Difficulty") +
  ylab("Discrimination") +
#  labs(colour = "Category") +
  labs(title = paste(language, scale)) +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "grey"),
    legend.position = "top"
  )
```

## Ability vs raw scores

We compute ability estimates ($\theta$'s) for children from the sample.

**NB:** For some datasets (models) the computation does not converge: should be looked into
(does it have something to do with the chosen MAP method?).

```{r ability}
scores <- cbind(scores, fscores(mod, method = "MAP", full.scores.SE = TRUE))
```

We plot estimated abilities ($\theta$'s) against raw scores (sums of checked items).

```{r ability_scores_plot}
ggplot(scores, aes(score, F1)) +
  geom_point() +
  ylab(expression(theta)) +
  xlab("Raw score") +
  labs(title = paste(language, scale)) +
#  ylim(-4, 4) +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "lightgrey"),
    text = element_text(size=16)
  )
```

The scatterplot above shows that IRT estimates get "normalised",
compared to raw scores.
The same can be seen comparing the following histograms.

```{r ability_scores_histograms}
# Should be in two separate chunks perhaps?
hist(scores$score, xlab = "Raw score", main = paste(language, scale))
hist(scores$F1, xlab = expression(theta), main = paste(language, scale))
```


## Ability vs SE

We plot ability estimates against their standard errors.

```{r ability_se_plot, fig.width=12, fig.height=8}
ggplot(scores, aes(F1, SE_F1)) +
  geom_point(alpha = 0.3) +
  xlab(expression(theta)) +
  ylab(expression(SE[theta])) +
  labs(title = paste(language, scale)) +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "lightgrey"),
    text = element_text(size=16),
    axis.title = element_text(size=20)
  )
```

The scatterplot above shows increasing errors at both extremes of the ability level
(perhaps due to floor and ceiling effects).
We also plot the distribution of SEs.

```{r se_histogram, fig.width=12, fig.height=8}
hist(scores$SE_F1, main = paste(language, scale), xlab = expression(SE[theta]))
```

We can export the data for crosslinguistic plotting/analysing.

```{r ability_se_export, eval=FALSE}
write_csv(scores, paste(filename_prefix, "thetas.csv", sep="_"))
```


## SE and ability level vs age

The graph below shows distribution of standard errors by age.

```{r se_age_plot}
ggplot(data.frame(age = as.factor(scores$age), SE = scores$SE_F1), aes(age, SE)) +
  geom_boxplot() +
  xlab("Age (months)") +
  ylab(expression(SE[theta])) +
  labs(title = paste(language, scale)) +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "lightgrey"),
    text = element_text(size=16)
  )
```

Now we plot the distribution of raw scores by age...

```{r score_age_plot}
ggplot(scores) +
  geom_point(aes(age, score)) +
  xlab("Age (months)") +
  ylab("Raw score") +
  labs(title = paste(language, scale))
```

...and the distribution of ability by age with colour indicating the ability's SE.

```{r ability_age_plot}
ggplot(scores) +
  geom_point(aes(age, F1, col = SE_F1)) +
  scale_fill_gradient(low="green", high="red", aesthetics = "col") +
  xlab("Age (months)") +
  ylab(expression(theta)) +
  labs(title = paste(language, scale), col = expression(SE[theta]))
#  ylim(-4, 4)
```


# Simulations

```{r prepare_sim}
params_mirt <- as.data.frame(coef(mod, simplify = TRUE)$items)[1:2]
mo <- generate.mirt_object(params_mirt, '2PL')
cl <- makeCluster(detectCores())
```

At the moment the simulations we run help us to decide CAT stopping criteria,
max. $SE_{\theta}$ accepted and max. number of items allowed. We first observe
how adding items decreases the mean $SE_{\theta}$ and we identify the $SE_{\theta}$
value, for which adding another item doesn't result in a *substantial drop* anymore.
Then we run simulations with that value as a stopping crietrion to see
how many items are needed to reach this criterion for a *reasonably high*
proportion of administrations.

```{r stop_criteria_parameters, echo=TRUE}
SE_delta <- 0.001 # non-substantial drop in mean SE
q_fail <- .2 # reasonably low proportion of administrations not reaching desired SE
```


## Whole pool for everyone

We simulate CAT administrations without any stopping criteria (and save the simulation
results to "`r paste(filename_prefix, "sim_all.RData", sep = "_")`" for later reuse).

```{r sim_all, message=TRUE}
list.files(path = data_dir, pattern = "sim_all\\.RData$", full.names = TRUE) |>
  str_match(paste(filename_prefix, "sim_all\\.RData$", sep = "_")) |>
  na.omit() -> sim_all_file
if(length(sim_all_file)) {
  load(sim_all_file)
} else if(isTRUE(getOption("knitr.in.progress"))) {
  stop(paste("Simulations might take a long time to compute.",
             "Run the code interactively in RStudio to see progress.",
             "The results will be saved to a file and you can try knitting again afterwards."))
} else {
  message(paste("Starting simulation with no STOP criteria @", Sys.time()))
  results_all <- mirtCAT(mo = mo, method = "MAP", criteria = "MI", start_item = "MI", cl = cl,
                         local_pattern = responses, design = list(min_items = ncol(responses)),
                         progress = TRUE)
  flush.console()
  save(results_all, file = paste(filename_prefix, "sim_all.RData", sep = "_"))
  message(paste("Finished @ ", Sys.time()))
  beep()
}
report_sim_results(results_all, scores, nrow(cdi))
```

Based on the whole pool simulation we can examine decreasing average $SE_{\theta}$ with each item.

```{r mean_se}
SE_history <- drop(simplify2array(map(results_all, "thetas_SE_history")))
meanSE_item <- as.data.frame(rowMeans(SE_history))
meanSE_item$item <- row.names(meanSE_item)
colnames(meanSE_item) <- c("mean_SE", "test_length")
meanSE_item$test_length <- as.numeric(meanSE_item$test_length)
```

We want to find the first value for which the difference between it and the next one
is less than `r SE_delta`.

```{r se_threshold}
se_plateau <- meanSE_item %>%
  mutate(diff = c(NA, diff(mean_SE))) %>%
  filter(!is.na(diff)) %>%
  filter(abs(diff) < SE_delta) %>%
  slice(1)
se_threshold <- se_plateau$mean_SE
se_plateau
```

We plot the average $SE_{\theta}$ by each consecutive item
(max. 100: highly unlikely not to reach the plateau by then)
and add red dashed lines marking the above.

```{r se_length_plot}
ggplot(meanSE_item, aes(x = test_length, y = mean_SE)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = se_threshold, linetype = "dashed", colour = "red") +
  geom_vline(xintercept = se_plateau$test_length, linetype = "dashed", colour = "red") +
  labs(title = paste(language, scale),
       x = "Test length",
       y = expression("Mean" ~ SE[theta])) +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_line(colour = "lightgrey"),
    text = element_text(size=16)
  ) + 
  scale_y_continuous(breaks = c(0.1, 0.15, 0.2, 0.25, 0.5, 0.75, 1)) + 
  scale_x_continuous(breaks = seq(0, nrow(params_mirt), by=10), limits = c(1, min(100, nrow(params_mirt))))
```

We can export the data for crosslinguistic plotting/analysing.

```{r mean_se_export, eval=FALSE}
write_csv(meanSE_item, paste(filename_prefix, "meanSE.csv", sep = "_"))
```


## $SE_{\theta} <$ `r ceiling(se_threshold * 1000) / 1000` stop criterion

We simulate CAT again, this time stopping each administration when its
$SE_{\theta}$ reaches `r ceiling(se_threshold * 1000) / 1000`.

```{r sim_se, message=TRUE}
se_precision <- nchar(strsplit(as.character(SE_delta), "\\.")[[1]][2])
se_string <- substr(str_split_i(se_threshold, "\\.", i=2), 0, se_precision)
list.files(path = data_dir, pattern = paste0("sim_", se_string, "\\.RData$"), full.names = TRUE) |>
  str_match(paste0(filename_prefix, "_sim_", se_string, "\\.RData$")) |>
  na.omit() -> sim_se_file
if(length(sim_se_file)) {
  load(sim_se_file)
} else if(isTRUE(getOption("knitr.in.progress"))) {
  stop(paste("Simulations might take a long time to compute.",
             "Run the code interactively in RStudio to see progress.",
             "The results will be saved to a file and you can try knitting again afterwards."))
} else {
  results <- sim_se(mo, responses, cl, se_threshold,
                    file = paste0(filename_prefix, "_sim_", se_string, ".RData"))
}
report_sim_results(results, scores, nrow(cdi))
```

We want to see the distribution of the number of items needed to reach it
(and the proportion of administrations that will not reach it at all).

*__NB:__ If the histogram below doesn't look ok, here are some settings*
*you can modify.*

```{r plot_length_hist_settings, echo=TRUE}
bin_width <- 10
pfs <-  6   # font size for percentages above bars
xfs <-  14  # font size for bin labels on X axis
```


```{r plot_length_hist}
plot_length(results, paste(language, scale), se_threshold, dim(responses), bin_width, pfs, xfs)
```

We want to find the number of items needed to leave only `r q_fail` of all
administrations not reaching $SE_{\theta} <$ `r ceiling(se_threshold * 1000) / 1000`.

```{r length_threshold}
sim_length_distro_q(results, q_fail)
```


# Config files preparation

For the CAT app to work we need two files, `items.csv` with parameter estimates and
a question for each item, and `startThetas.csv` with $\theta_0$, either a single value
or different values for different (combinations of) demographic (age and/or sex) variables.

$\theta_0$ values may be "corrected" (i.e., decreased by a set constant) to make it more
likely for first items to be answered positively.

"Questions" for individual items may be used to display for each item its category, but
they may also be more specific (which is important for gestures but might be helpful
to keep track of the comprehension/production distinction as well).

The code chunk below, if run, creates two files:
"`r paste(filename_prefix, "items.csv", sep = "_")`" and
"`r paste(filename_prefix, "startThetas.csv", sep = "_")`".

```{r config_files, eval=FALSE, echo=TRUE}
start_thetas <- get_start_thetas(scores, "age", "sex", correction=.5)[[1]]
prepare_config_files(params_mirt, cdi$item_definition, fscores_aggr,
                     question = "CATEGORY LABEL here", group = subscale, id = cdi$item_id,
                     file1 = paste(filename_prefix, "items.csv", sep = "_"),
                     file2 = paste(filename_prefix, "startThetas.csv", sep = "_"))
```



